---
title: 'Assignment 2: Data visualization'
author: "Dorottya Csala"
output: html_document
editor_options: 
  chunk_output_type: console
---

You will have to create 3 plots based on the datasets and instructions detailed below. You will find the plots themeselves in the `assignments/assignment_2_plots`. Your task is to write the code that will reproduce the plots as closely as possible.

# Skills needed to solve this assignment

-   Using R and RStudio, reading data
-   Reporting using RMarkdown
-   Using Git and Github (for submitting the task)
-   Data manipulation (e.g. dplyr, tidyr), and working with factors (forcats)
-   Data visuzlization (ggplot2)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(tidytuesdayR)
library(readr)
library(ggplot2)

expeditions <- read_csv("expeditions.csv")

head(expeditions)
colnames(expeditions)

expeditions_peaks <- expeditions %>%
  mutate(most_peaks= fct_lump(peak_name, n= 15)) %>%
  filter(most_peaks != "Other") %>%
  count(most_peaks, season) %>%
  group_by(most_peaks) %>%
  mutate(total = sum(n)) %>%
  ungroup() %>%
  mutate(most_peaks= fct_reorder(most_peaks, total, .desc= FALSE))
  
print(expeditions_peaks)


ggplot(expeditions_peaks, aes(x= n, y= most_peaks)) + geom_col(aes(fill=season))  + labs(
  title= "The 15 most popular peaks stacked by season of expedition",
  x= "Number of expeditions", y= "", fill="Season") + scale_fill_viridis_d(option="D") + theme_light() + theme(legend.position="bottom")




```

## Task 1: Climbing expeditions

The 2020-09-22 TidyTueday datasets are about climbing expeditions. From the three datasets, use the "expeditions". Reproduce the plot below! Notice a few things:

-   Use `forcats::fct_lump()` to get the 15 most frequent peaks, and drop the "Other" category.
-   The bars are ordered by the sum of all expeditions (use `fct_reorder()`).
-   The bar colors use the viridis palette and light theme.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(readr)
library(ggplot2)
library(scales)

phd_by_field <- read.csv("phd_by_field.csv")
head(phd_by_field)

broad_field_phds <- aggregate(n_phds ~ year + broad_field,data= phd_by_field, FUN= sum) 
  
print(broad_field_phds)

ggplot(broad_field_phds, aes(x= year, y= n_phds, color= broad_field)) + geom_line(linewidth= 1.2) + scale_x_continuous(breaks= pretty_breaks(n= 5)) + scale_y_continuous(labels= comma_format())+ labs(title= "Number of awarded Ph.D.-s in the US by year", x= "", y= "", color= "Broad field") + scale_color_brewer(palette="Dark2") + theme_minimal()
```

## Task 2: PhDs awarded

The 2019-02-19 TidyTueday dataset is about phd-s awarded by year and field. There is only one dataset, it is called `phd_by_field`. Reproduce the plot below!

Notes:

-   First you have to aggregate the data to count all phd-s by broad fields.
-   To make the x axis breaks pretty, use `scales::pretty_breaks()`, to make the y axis labels comma formatted, use `scales::comma_format()`.
-   The line size is 1.2, the colors are from the brewer "Dark2" palette. The theme is set to minimal.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(dplyr)
library(readr)
library(ggplot2)
library(scales)

commute <- read.csv("commute.csv")
commutes_by_state <- commute %>%
  group_by(state, mode, state_region, state_abb) %>%
  summarise(n = sum(n, na.rm = TRUE)) %>%
  ungroup() 
commutes_wide <- commutes_by_state %>%
  pivot_wider(names_from = mode,
              values_from = n,
              values_fill = 0) %>%
  rename(bike = `Bike`, walk = `Walk`) 
colnames(commutes_wide)

states_labels <- c("WY", "VT", "ND", "AK", "DE", "NH", "ME", "MT", "ID", "RI", "NE", "KS", "LA", "IA", "CT", "UT", "IN", "MI", "MN", "AZ", "OR", "FL", "WA", "MA", "IL", "OH", "NJ", "OA", "NY", "CA")

ggplot(commutes_wide, aes(x = walk, y = bike, colour = state_region, label= state_abb)) + 
  geom_point(size = 2) +   geom_text(data= subset(commutes_wide, state_abb %in% states_labels), hjust = 0.5, size = 5, color= "black") +
  scale_x_log10(labels = comma_format()) +
  scale_y_log10(labels = comma_format()) +
  labs(
    title = "Title number of people walking vs. biking to work in each USA state",
    x = "Number of ppl walking to work (log N)",
    y = "Number of ppl biking to work (log N)", color= "State region"
  ) +
  theme_light()


```

## Task 3: Commute in the US

The 2019-11-05 TidyTueday dataset is about commuting to work in each city in the US by bike or on foot. There is only one dataset, it is called `commute`. Reproduce the plot below!

Notes:

-   First you have to aggregate the data to count all commutes by state.
-   Both axis scales are log transformed and the labels comma formatted, using `scales::comma_format()`
-   The point size is 2, . The theme is set to light.

```{r}

```
